---
title: "Enough heatmaps! Easy visualisations of correlations"
author: "Makowski, D. and Lüdecke, D."
date: 2019-03-08
categories: ["R", "correlation"]
tags: ["R", "easystats", "correlation", "plot", "graph", "ggraph", "corrplot", "radar chart", "spider plot"]
summary: This post shows how to easily create correlation graph plots with the use of the correlation package.
editor_options: 
  chunk_output_type: console
---



Recently, [Daniel](https://github.com/strengejacke) (author of great packages such as [sjPlot](https://github.com/strengejacke/sjPlot) and [ggeffects](https://github.com/strengejacke/ggeffects)) and I (from the [psycho](https://github.com/neuropsychology/psycho.R) package), decided to unite our forces to provide **flexible, stable and lightweight** tools for facilitating data analysis. The [easystats](https://github.com/easystats) project, altough in very active developpement, already contains some useful functions.

One of such can help creating **pretty correlation graphs and plots**. [correlation](https://github.com/easystats/correlation) is a small package designed to compute correlations in a format that can easily be fed into plotting packages.

## Packages

As the [**correlation**](https://github.com/easystats/correlation) package is not (yet) on CRAN, you need to install it from github:

```{r message=FALSE, warning=FALSE, eval=FALSE}
install.packages("devtools")
devtools::install_github("easystats/correlation")
library(correlation)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(correlation)
```


Let's also install [**report**](https://github.com/easystats/report), another package from the *easystats* project.

```{r message=FALSE, warning=FALSE, eval=FALSE}
install.packages("devtools")
devtools::install_github("easystats/report")
library(report)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(report)
```

These other packages will be used for data and plotting.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(forcats)
library(psych)
library(ggplot2)
library(tidygraph)
library(ggraph)
library(plotly)
```




## The data

From `psych`'s description: 

> State Anxiety was measured two-three times in 11 studies at the Personality-Motivation-Cognition laboratory. Here are item responses for 11 studies (9 repeated twice, 2 repeated three times). In all studies, the first occasion was before a manipulation. In some studies, caffeine, or movies or incentives were then given to some of the participants before the second and third STAI was given.

```{r message=FALSE, warning=FALSE}
# Load the data
df <- psych::sai %>% 
  filter(time < 4) %>% 
  mutate(time = as.factor(time),
         id = as.character(id)) %>% 
  select(time, relaxed, joyful, confident, anxious, worrying, upset)

# Report the data
report(df)
```

## Correlation

The main function of the *correlation* package is [`correlation()`](https://easystats.github.io/correlation/reference/correlation.html), with which you can tweak the correlation parameters (partial, semi-partial, Bayesian...).

```{r message=FALSE, warning=FALSE, eval=FALSE}
# Do the correlation
cor <- correlation(df)

# Examine the output
head(cor)
```
```{r message=FALSE, warning=FALSE, echo=FALSE}
cor <- correlation(df)

knitr::kable(head(cor), digits=2)
```

Contrary to the majority of packages, the `correlation()` function does not return a nice matrix, but rather a dataframe with pairwise correlations as rows. Altough this format will be useful for further plotting, we can always transform it to a **table** using the [`report`](https://github.com/easystats/report) package.


```{r message=FALSE, warning=FALSE, eval=FALSE}
cor %>% 
  report() %>% 
  to_table()
```
```{r message=FALSE, warning=FALSE, echo=FALSE}
cor %>% 
  report() %>% 
  to_table() %>% 
  knitr::kable()
```

## Heatmaps

Hetmaps can be nicely improved by removing the upper triangular using the `remove_triangular()` function.

```{r message=FALSE, warning=FALSE, eval=TRUE}
p <- cor %>% 
  remove_triangular() %>% 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  scale_fill_gradient2(low = "#d50000", high = "#00C853") +
  theme_minimal() +
  theme(axis.title = element_blank()) 
ggplotly(p)
```

## Bar plots

But enough with the heatmaps! There are other ways to present correlations. For instance, bar charts:

```{r message=FALSE, warning=FALSE, eval=TRUE}
p <- cor %>% 
  filter(Parameter1 != Parameter2) %>%
  ggplot(aes(x=Parameter1, y=r, fill=r)) +
  geom_bar(stat="identity") +
  scale_fill_gradient2(low = "#d50000", high = "#00C853") +
  theme_minimal() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Parameter2) +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(p)
```


## Radar chart

Or, if we want to get fancier:

```{r message=FALSE, warning=FALSE, eval=TRUE}
# Create a small convenience function
coord_radar <- function (theta = "x", start = 0, direction = 1) 
{
    theta <- match.arg(theta, c("x", "y"))
    r <- if (theta == "x") 
        "y"
    else "x"
    ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
        direction = sign(direction),
        is_linear = function(coord) TRUE)
}


p <- cor %>% 
  ggplot(aes(x=Parameter1, y=r, group=Parameter2, fill=Parameter2)) +
  geom_polygon(alpha=0.8, show.legend = FALSE) +
  theme_minimal() +
  theme(axis.title.x = element_blank()) +
  facet_wrap(~Parameter2) +
  coord_radar() 
p
```


## Graphs


However, one of the most interesting way of visualising correlations is through graphs. This can be done easily using the powerful [**ggraph**](https://github.com/thomasp85/ggraph) package.



```{r message=FALSE, warning=FALSE, eval=TRUE}
set.seed(999)
cor %>% 
  remove_triangular() %>% 
  as_tbl_graph() %>% 
  ggraph(layout = 'drl') +
  geom_edge_arc(aes(colour=r), curvature=0.1, edge_width=1) +
  geom_node_point(color="#607D8B", size=22) +
  geom_node_text(aes(label = name), colour="white") +
  scale_edge_color_gradient2(low = "#d50000", high = "#00C853") +
  theme_graph() +   
  scale_x_continuous(expand = expand_scale(c(.10, .10))) +
  scale_y_continuous(expand = expand_scale(c(.10, .10)))
```

## Interested by the easystats project?

The **easystats** project is very contributor-friendly, just share your ideas, opinions or suggestions on the [issues section](https://github.com/easystats/insight/issues).
