---
title: "Describe and understand Bayesian models and posteriors using bayestestR"
author: "Makowski, D. and Lüdecke, D."
date: 2019-04-12
categories: ["R", "bayestestR"]
tags: ["R", "easystats", "bayestestR", "posterior", "rstanarm", "brms"]
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
```




While the Bayesian framework is gaining popularity among scientists, leading to the growing popularity of packages to fit Bayesian models, such as [`rstanarm`](https://github.com/stan-dev/rstanarm) or [`brms`](https://github.com/paul-buerkner/brms). However, extracting summary indices from these models to report them in your manuscript can be quite rather challenging, especially for new users.

**Wecolme to bayestestR!**

## bayestestR

Recently, [we](https://github.com/orgs/easystats/people) have decided to collaborate around the new [easystats](https://github.com/easystats) project, a set of packages designed to make your life easier (currently very work in progress). As we are also Bayesian enthusiasts, we decided to focus on the development of a package devoted to Bayesian post-processing.

We are happy to present the new [**`bayestestR`**](https://github.com/easystats/bayestestR) R package.

The goal of bayestestR is to provide lightweight tools to help **processing and understanding Bayesian models and posterior distributions**. It includes several functions used to report and characterise them, such as [Highest Density Interval (`hdi`)](https://github.com/easystats/bayestestR#highest-density-interval-hdi---the-credible-interval-ci), the [highest Maximum A Posteriori (`MAP`)](https://github.com/easystats/bayestestR#map-estimate) or functions to find a suitable [ROPE range](https://github.com/easystats/bayestestR#find-ropes-appropriate-range), compute the [ROPE percentage](https://github.com/easystats/bayestestR#rope) or do an [Equivalence Test](https://github.com/easystats/bayestestR#equivalence-test). It also includes more exploratory indices, such as the [Probability of Direction (`pd`)](https://github.com/easystats/bayestestR#probability-of-direction-pd), a [numeric equivalent](https://easystats.github.io/bayestestR/articles/guidelines.html) of the Frequentist *p* value.

`bayestestR` can be installed as follows:

```{r, eval = FALSE, message=FALSE, warning=FALSE}
install.packages("bayestestR")  # Install the package
library(bayestestR)  # Load it
```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(bayestestR) 
```

- **Don't forget to check out the** [**documentation here**](https://github.com/easystats/bayestestR#documentation)**!**


## Future

There is definitely room for improvement, and some new exiting features are already planned (BayesFactor objects support, better visualisation methods, etc.). Feel free to let us know how we could further improve this package!

To conclude note that *easystats* is a new project in active development, do not hesitate to contact us if **you want to get involved :)**

- **Check out our other blog posts** [***here***](https://easystats.github.io/blog/posts/)!
