---
title: "How to easily generate a perfectly normal distribution"
author: "easystats"
date: 2019-04-24
categories: ["R", "bayestestR"]
tags: ["R", "easystats", "bayestestR", "rnorm", "normal", "gaussian"]
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
```



Many times, for instance when teaching, I needed to quickly and simply generate a **perfectly normally distributed sample** to illustrate or show some of its characteristics.

This is now very easy to do with the new [`bayestestR`](https://github.com/easystats/bayestestR) package, which includes the [`rnorm_perfect`](https://easystats.github.io/bayestestR/reference/rnorm_perfect.html) function. This function is very similar to the classic `rnorm` (same arguments), with the difference that the generated sample is *perfectly* normal.


## Example

`bayestestR` can be installed as follows:

```{r, eval = FALSE, message=FALSE, warning=FALSE}
install.packages("bayestestR")  # Install the package
library(bayestestR)  # Load it
```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(bayestestR)
```
```{r message=FALSE, warning=FALSE,}
# Generate a perfect sample
x <- rnorm_perfect(n = 100, mean = 0, sd = 1)

# Visualise it
library(tidyverse)

x %>% 
  density() %>%  # Compute density function
  as.data.frame() %>% 
  ggplot(aes(x=x, y=y)) +
  geom_line()
```

We can also easily color some of the parts of the curve, for instance, the observations lying beyond +2 standard deviations.

```{r message=FALSE, warning=FALSE,}
x %>% 
  density() %>%  # Compute density function
  as.data.frame() %>% 
  mutate(outlier = ifelse(x > 2, "Extreme", "Not extreme")) %>% 
  ggplot(aes(x=x, y=y, fill=outlier)) +
  geom_ribbon(aes(ymin=0, ymax=y)) +
  theme_classic()
```

## bayestestR and easystats


More details about `bayestestR`'s features are comming soon, stay tuned ;)

- **Don't forget to check out the** [**documentation here**](https://github.com/easystats/bayestestR#documentation) **for more!**


Feel free to let us know how we could further improve this package! Also, note that [*easystats*](https://github.com/easystats/easystats), the project supporting `bayestestR` is in active development. Thus, do not hesitate to contact us if **you want to get involved :)**

- **Check out our other blog posts** [***here***](https://easystats.github.io/blog/posts/)!
