---
title: "A perfectly normally distributed sample: Another post?"
author: "easystats"
date: 2019-04-24
categories: ["R", "bayestestR"]
tags: ["R", "easystats", "bayestestR", "rnorm", "normal", "gaussian"]
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
```


Recently, we published [a post](https://easystats.github.io/blog/posts/bayestestr_rnorm_perfect/) presenting a small convenience function, from the [bayestestR package](https://github.com/easystats/bayestestR). This function, named `rnorm_perfect`, generated an empirical distribution (*i.e.*, vector of values) as close as possible to a desired distribution, in this case the *normal* distribution. 

It worked like this:

```{r message=FALSE, warning=FALSE, include=FALSE}
library(bayestestR)
```
```{r message=FALSE, warning=FALSE}
# Generate a perfect sample
x <- rnorm_perfect(n = 100, mean = 0, sd = 1)

# Visualise it
library(tidyverse)

x %>% 
  density() %>%  # Compute density function
  as.data.frame() %>% 
  ggplot(aes(x=x, y=y)) +
  geom_line()
```


This function's name was initially chosen because of the argument-wise proximity of this function with its random counterpart, `rnorm` (*i.e.*, it has the same arguments, the same order and the same output type). However, we [quickly realised](https://github.com/easystats/bayestestR/issues/86) this this name was not perfect, as indeed the obtained distribution was not *random* (and the `r` in `rnorm` stands for *random*).

Thus, after discussion, we decided to change it to `distribution_normal` (INSERT LINK). And guess what, we also added other types of "perfect" distributions:



```{r message=FALSE, warning=FALSE}
# Generate a perfect sample
x <- distribution_poisson(n = 100, 1.5)

x %>% 
  estimate_density("logspline") %>% 
  as.data.frame() %>% 
  ggplot(aes(x=x, y=y)) +
  geom_line()
```

A few days before this update, we received a small email asking:

> Hi, I saw your blog post and wonder how you define a perfectly normal distribution.

We responded, as honestly as we could:

> that's actually a good question. I would say an empirical sample having characteristics as close as possible to a canonic Gaussian distribution.

And just now, after an (ironic?) *"Thanks, most helpful!"*, [this blog post](https://xianblog.wordpress.com/2019/05/09/a-perfectly-normally-distributed-sample/) is published, which emphasize on the irrelevance of the `r` prefix in the function name. A great timing, as we wanted to share the update that we made! 

Note that the `rnorm_perfect` name will be deprecated in the next version, and ultimately removed to avoid further confusion :)


## Any other suggestions?

Don't forget, the [easystats project](https://github.com/easystats) (that includes bayestestR) is very open to contributions and suggestions!

Thus, if you ever plan on publishing personal correspondence with the response to a falsely naive question, please remember that you can always [make suggestions](https://github.com/easystats/bayestestR/issues) and contribute to the package, to *actually help* improving it, with the spirit of collaboration, and in the tradition of open science ;)


## Get involved


Feel free to let us know how we could further improve this package! Also, note that [*easystats*](https://github.com/easystats/easystats), the project supporting `bayestestR` is in active development. Thus, do not hesitate to contact us if **you want to get involved :)**

- **Check out our other blog posts** [***here***](https://easystats.github.io/blog/posts/)!
