---
title: "The p-direction: A Bayesian equivalent of the p-value?"
author: "easystats"
date: 2020-02-26
categories: ["R", "bayestestR"]
tags: ["R", "easystats", "bayestestR", "p-value", "pd", "p-direction"]
editor_options: 
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
set.seed(333)
```



The Bayesian framework is powerful and allows for an incredible amount of flexibility and control over your analysis. That being said, newcomers often struggle with a lot of new concepts and tools and could benefit from some **familiar grounding**. And the *p*-value is a very familiar index (although paradoxically often misunderstood, but that's another topic).

Is there an equivalent of the *p*-value? Well, depends on what "equivalent" means. Some might argue that the [**Bayes factor**](https://easystats.github.io/bayestestR/articles/bayes_factors.html) is some sort of equivalent, i.e., a value that can be used for decisions and interpretation of results. [Some others](https://www.youtube.com/watch?v=Ip8Ci5KUVRc&t=408s) would suggest that the [MAP-based *p*-value](https://easystats.github.io/bayestestR/reference/p_map.html) is another equivalent.

Based on a simulation study [(Makowski et al., 2019)](https://www.frontiersin.org/articles/10.3389/fpsyg.2019.02767/full), we think that the [**probability of direction**](https://easystats.github.io/bayestestR/articles/probability_of_direction.html) (*p*-direction, or *pd*) is the closest ***statistical* equivalent to the *p*-value***. The *statistical* is important here, simply meaning that the two indices are strongly correlated. That being said, they are **not *conceptually* equivalent** (as we argue in the paper, the *pd* is an index of effect *existence*, rather than *significance*).

Here's a short example.


## Frequentist regression

First, you can install (or update) the necessary packages by running the following (it's important that the *insight* package version must be >= 0.8.1):

```{r message=FALSE, warning=FALSE}
install.packages(c("insight", "bayestestR", "parameters"))
```

Let's start by running a simple linear regression and displaying its result with the [**parameters**](https://easystats.github.io/parameters/) package.



```{r message=FALSE, warning=FALSE}
library(parameters)

model <- lm(disp ~ carb, data = mtcars)
parameters(model)
```

The *p*-value of the linear relationship between the two variable is of *.025*. What does a Bayesian analysis tells us?


## Bayesian regression (with flat priors)

As you might know, a Bayesian analysis is close to a maximum likelihood analysis (the typical frequentist paradigm) when no information is given by the prior (and the result is only driven by the data). This is the case with **flat priors**, that give equivalent likelihood to each and every one of your wildest dreams (see the [*How to Specify Flat Priors (and why you typically shouldn’t)*](https://cran.r-project.org/web/packages/rstanarm/vignettes/priors.html) section).

Let's fit the same regression as above within a Bayesian framework with a flat prior. This will take a while to run, because we will specify a lot of iterations (`8*100000`) to increase the precision and stability of our results.

```{r message=FALSE, warning=FALSE, results='hide'}
library(bayestestR)
library(rstanarm)

model <-
  stan_glm(
    disp ~ carb,
    data = mtcars,
    priors = NULL,
    prior_intercept = NULL,
    chains = 10,
    iter = 10000,
    seed = 333
  )
```
```{r message=FALSE, warning=FALSE}
parameters(model)
```

It tells us that the *p*-direction is of 98.69\% (i.e., `0.9869`). We can quickly visualize its meaning as follows (with the [**see**](https://github.com/easystats/see) package):

```{r message=FALSE, warning=FALSE}
library(see)

plot(p_direction(model))
```


## From *p*-direction to *p*-value

We can convert this value to a *p*-value using the following function:

```{r message=FALSE, warning=FALSE}
pd_to_p(0.9869)
```

As we can see, we are not far from the frequentist *p*-value! 

But again, we need to underline that the *p*-direction has a **different meaning and interpretation**. It refers to the *probability that the effect is positive or negative (depending on the median's sign)*. But like the *p*-value, it cannot either be used to **support a lack of an effect** (for that, [ROPE-based indices](https://easystats.github.io/bayestestR/articles/region_of_practical_equivalence.html) or [Bayes factors](https://easystats.github.io/bayestestR/articles/bayes_factors.html) might be more appropriate). 

Moreover, when using informative priors centred at 0, a Bayesian analysis will always lead to "less significant" effects, as the prior will pull the posterior towards 0. This is a natural way of penalizing results, which **is a good thing**.

In conclusion, make sure you understand the indices you use (for instance by [**checking-out our gentle intro to Bayesian analysis**](https://easystats.github.io/bayestestR/articles/bayestestR.html))!



## Get Involved

*easystats* is a new project in active development, looking for contributors and supporters. Thus, do not hesitate to contact us if **you want to get involved :)**

- **Check out our other blog posts** [***here***](https://easystats.github.io/blog/posts/)!

## Stay tuned 

To be updated about the *upcoming features* and cool R or data science stuff, you can **follow the packages on GitHub** (click on one [of the easystats package](https://github.com/easystats)) and then on the **Watch** button on the top right corner) as well as the **easystats team on twitter and online**:

- [\@easystats4u](https://twitter.com/easystats4u)
- [\@strengejacke](https://twitter.com/strengejacke)
- [\@DominiqueMakowski](https://twitter.com/Dom_Makowski)
- [\@mattansb](https://twitter.com/mattansb)
- [\@IndrajeetPatil](https://twitter.com/patilindrajeets)
- [\@humanfactors](https://twitter.com/humanfactorsio)
- [\@lindeloev](https://twitter.com/jonaslindeloev)
- [\@jacob-long](https://twitter.com/jacobandrewlong)
- [\@pdwaggoner](https://pdwaggoner.github.io/)
- [\@LeoDutriaux](https://twitter.com/LDtrx)
