---
title: "A unified syntax for accessing models' information"
author: "Lüdecke, D. and Makowski, D."
date: 2019-04-03
categories: ["R", "insight"]
tags: ["R", "easystats", "insight", "access", "models", "data"]
---


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
options(knitr.kable.NA = '')
```




The richness and variety of packages for building and fitting statistical models in R is absolutely astonishing and contributes to the language's popularity. However, **this diversity makes it hard for developpers** that want to create tools that work with different types of models. Indeed, the way to access models' internal information (such as **parameters names**, **formulae**, **data**, etc.) is **not unified**, forcing the developers to spend some time figuring out how to do it for each model type.

**This time is over!**

## Insight

Recently, we ([DL](https://github.com/strengejacke) and [DM](https://github.com/DominiqueMakowski)) have decided to collaborate around the new [easystats](https://github.com/easystats) project, a set of packages designed to make your life easier (currently a very work in progress). However, in order to create these packages and functions, **we needed a basis**, a stable cornerstone, that would allow the unified way of accessing models information. 

And [**`insight`**](https://github.com/easystats/insight) was born.

The goal of insight is to provide tools to help an **easy, intuitive and consistent accesss** to information contained in various models. Indeed, although there are generic functions to get information and data from models, many modelling-functions from different packages do not provide such methods to access these information. The insight package aims at closing this gap by providing functions that work for (almost) any model.

`insight` can be installed as follows:

```{r, eval = FALSE, message=FALSE, warning=FALSE}
install.packages("insight")  # Install from CRAN
library(insight)  # Load the package
```
```{r message=FALSE, warning=FALSE, include=FALSE}
library(insight) 
```

## Example

Let's see how it works on a very simple regression model:

```{r message=FALSE, warning=FALSE}
model <- lm(Sepal.Length ~ Species, data=iris)
```

- Find the **parameters**:

```{r message=FALSE, warning=FALSE, comment=">"}
find_parameters(model)
```

- Find the **outcome's name**:

```{r message=FALSE, warning=FALSE, comment=">"}
find_response(model)
```

- Find the **formula**:

```{r message=FALSE, warning=FALSE, comment=">"}
find_formula(model)
```

- Find the **variables in the formula**:

```{r message=FALSE, warning=FALSE, comment=">"}
find_variables(model)
```

- Find the **algorithm**:

```{r message=FALSE, warning=FALSE, comment=">"}
find_algorithm(model)
```


Moreover, `insight` also includes function to deal with **Bayesian models** (`get_priors()`) and **mixed models** (`find_random()`).

`insight` works on a high number of models (see the [**list here**](https://github.com/easystats/insight/blob/master/README.md#list-of-supported-packages-and-models)), and **continue to grow thanks to your suggestions**! As *easystats* is a new project in active development, do not hesitate to contact us if **you want to get involved :)**

- **Check out our other blog posts** [***here***](https://easystats.github.io/blog/posts/)!
